Glyph-maps for visualizing climate data and models

### 2011-08-31 paper.tex
computation section

A big advantage of glyph-maps is that they can be generated with existing graphics software after performing a simple pre-processing step described in this section. This makes them immediately accessible to a wide audience. The graphics in this paper were created in R \citep{R} using ggplot2 \citep{me:ggplot2,wickham:2007d} and \citep{me:plyr}, but glyph-maps are readily implemented in any statistical programming environment.

The pre-processing step is simple, once you notice that the position of graphical elements is determined by a major and minor axis in each direction ($x$ and $y$). For maps, the major axes are latitude and longitude, and the minor axes are typically time on the $x$-axis and some measurement or model prediction on the $y$-axis. The final locations are a simple linear combination.  After rescaling the minor axis to have range $[0, 1]$, the coordinates are given by

@@

Computing the coordinates for star glyphs is achieved by using polar coordinates of $y_{\amin}$, and $x_{\amin}$,  $r$, the radius scaled to be between 0 and 1, and $\theta$, the angle. Then the glyph coordinates are given by

@@

Figure~\ref{fig:cycle} shows monthly temperature effects as time (Cartesian coordinates) and star (polar coordinates) glyphs. In this case, the star glyphs are not as effective as the linear glyphs: perhaps because of the absence of cyclical pattern.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=2in]{month-cartesian}
  \includegraphics[width=2in]{month-polar}

@@

>>>>

A big advantage of glyph-maps is that they can be generated with existing graphics software after performing a simple pre-processing step, making them easily accessible to a wide audience. The pre-processing step is simple, once you notice that the position of graphical elements is determined by a major and minor axis in each direction ($x$ and $y$). For maps, the major axes are latitude and longitude, and the minor axes are typically time and some measurement or model prediction. In addition, the minor axes are rescaled to have range $[0, 1]$. The final data values are a simple linear combination, as given by

@@

Computing the coordinates for star glyphs is achieved by using polar coordinates,  $r=y_{\amin}$, and $\theta=x_{\amin}$, both scaled to be between 0 and 1. Then the glyph coordinates are given by

@@

Figure~\ref{fig:cycle} shows average monthly temperature as time (Cartesian coordinates) and star (polar coordinates) glyphs, for a subset of the spatial region. In this case, the star glyphs are not as effective as the linear glyphs: perhaps because of the absence of cyclical pattern.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=3in]{month-cartesian}
  \includegraphics[width=3in]{month-polar}

@@

The graphics in this paper were created in R \citep{R} using {\tt ggplot2} \citep{me:ggplot2} and {\tt plyr} \citep{me:plyr}.
